Analytical Model for the Current Density in the Electrochemical Synthesis of Porous Silicon Structures with a Lateral Gradient

### 2020-12-29 paper.tex
Change reflectance plots by migrad fits
Include position uncertainty
Advance in the reply to the referee

  non-homogenous multilayered devices with lateral gradients for

@@

\section{Introduction}
\label{sec:introduction}
Surfaces with a gradient in the refractive index (GRIN) are of
importance as they allow the engineering and tuning of the optical
phase along their surface, yielding applications such as flat
lenses. Furthermore, there are important biological applications
which combine topographical features with index of refraction
gradients \cite{pierscionek2012gradient}. Preparation of structures
with gradients in other properties have also shown their
usefulness. As an example, the
roughness dependent  response of cells has been tested on
materials used for applications in medical implants
\cite{bai2020bioinspired}.
Gradients have also been employed for several
technological applications such as high-yield screening of catalysts,
sensing materials and biomolecules \cite{maier2007combinatorial, jayaraman2004construction,
potyrailo2008combinatorial}
Apart from some methods developed by S.E. Fosdick et al \cite{fosdick2010two},
%[JACS  2010, 132, 9226-9227],
synthesis of  Ag-Au alloy gradients on steel and chemical composition
gradients of CdS layers
on gold electrodes have also been obtained \cite{wang2019fabrication}.
Among electrochemical methods, compositional and doping density changes
in  conducting polymers have been used to produce gradients using Indium Tin
Oxide electrodes \cite{inagi2010bipolar}.
Electrochemically induced potential gradients have been shown to catalyze the
reactions and gradient doping of polymers \cite{qin2020bipolar, ishiguro2011gradient}.
As compared to the above mentioned techniques, a relatively economical,
fast and easy approach for the fabrication
of gradient materials has been the use  of asymmetrical
electrode configurations in the electrochemical synthesis of porous silicon \cite{sailor2012porous}.
This technique has the additional advantage of being compatible
with very large scale integration (VLSI) devices and integrable with microelectronics.

Although porous silicon (PS) initially stimulated the interest of the scientific
community due to its photoluminescence in the visible range at room
temperature, it has presently
been recognized as a multifaceted optical material due to its large surface
area, biocompatibility, ease of fabrication and tunable refractive index.
Applications based on porous silicon now cover various fields such as
chemical sensors and biosensors \cite{kumar2020porous},
microelectronics and micromechanical systems (MEMS) \cite{martinez2016dual},as well as a
range of optical \cite{estevez2014porous} and optoelectronic applications
\cite{ramadan2020fabrication}. Specifically,
the temporal variation of the current density results in a variation of
porosity along the depth, allowing the easy fabrication of different kinds
\cite{perez2018reflectivity} of 1D dielectric multilayered
structures.
Although the conventional fabrication of PS yields laterally
homogeneous samples, with invariant structural and optical properties
along the surface, another configuration where a Pt pin electrode is placed
perpendicular to the silicon wafer, produces porous
silicon structures with a lateral gradient in properties such as
thickness, pore size and density, and thus, refractive index and
optical thickness \cite{khung2008using, clements2011mesenchymal}.
The resulting porous surface can have pore sizes ranging from a few nanometers
to  few micrometers  \cite{wang2015screening}. The range of the dimensions of the pores and
the corresponding  thickness range on the same chip can be controlled by
adjusting the location of the electrode, the  anodizing current
and the etching time. Different
applications have been found for the resulting structures
\cite{collins2002determining}. For example, Sailor's group
\cite{sailor2012porous, collins2002determining} applied it
for the determination of protein size.
Additionally, a variation in pH of the solution could gate the
trapping/release of biomolecules at regions with different porosity, a
result potentially useful for drug delivery
applications given the biocompatible nature of PS.
A similar asymmetric electrode configuration was used to fabricate
multilayered optical filters with lateral a gradient
\cite{li2005painting} and
for developing ethanol sensors \cite{sailor2012porous}.
Additionally, it has been shown that the combination of thermal  oxidation  and
infiltration of TiO2  by atomic layer deposition allows the manufacture
of transparent GRIN
optical elements with a high refractive index contrast \cite{ocier2017tunable}.
Kruger et al. \cite{krueger2018electrochemical} used
a porous silicon 1D GRIN structure infiltrated with polymer to
synthesize flat microlens arrays.
Recently, J Wang et al. has shown the fabrication of
a miniature spectrometer with a PS based rugate filter using a
radial interfacial potential distribution \cite{wang2019fabrication}
On the other hand, the effect of different topographical
features present on the same chip, has also been
used to study the dependence of the adhesion of certain cells on the
surface topography \cite{khung2008using}.

@@

with that within an infinite fictitious empty space but for the point charge

@@

The electrostatic potential produced by an infinite array of
charges expressed as a sum in real space of Coulomb terms has ill
convergence properties. The potential can be obtained
using the Ewald summation technique
\cite{ewald1921calculation, belhadj1991molecular-dynamics,
  schmidt1997multipole, grzybowski2000ewald, toukmaji2000efficient,
  brodka2004ewald, campione2012ewald}, which yields
two rapidly converging series, one in real space and another one in reciprocal space.
Nevertheless, as we need the potential and the field to obtain the
current density on the surface of the sample on which  no
real nor image charges lie, we may perform a somewhat simpler planewise
summation, where we first obtain the potential and field produced by a
simple periodic lattice of charges using a fast convergent sum over
the two dimensional (2D)
reciprocal vectors \cite{nijboer1957calculation, nijboer1958internal,
  de_wette1965internal, watson1981madelung, grindlay1981k},
then sum over the charges that form the basis
of the 2D {\em crystal} plane, and finally sum over all the
planes.

We consider first a system made of identical {\em unit} charges
ocuppying the positions $\bm R$ of a 2D Bravais lattice $\{\bm R\}$,
which we will take as a rectangular lattice with lattice parameters
$2a$ and $2b$, and lying, for the time being, at the $z=0$ plane.
The potential $\phi$ it produces obeys
\begin{equation}
  \label{eq:poissonRed}
  \nabla^2\phi(\bm r)=-4\pi\sum_{\bm R}\delta^{(2)}(\bm r_\|-\bm
  R)\delta^{(1)}(z)=-\frac{4\pi}{A}\sum_{\bm G}e^{i\bm G\cdot\bm r_\|}\delta^{(1)}(z),
\end{equation}
where $\bm r_\|$ is the projection of the observation position $\bm
r=(x,y,z)$ onto the $xy$ plane, $\{\bm G\}$ is the 2D
reciprocal lattice defined through $e^{i\bm G\cdot\bm R}=1$ for all
$\bm R$ and $\bm G$, $A=4ab$ is the area of the unit cell
and we used the Fourier representation of the 2D
periodically repeated delta function. We introduce a 2D Fourier representation for
the potential
\begin{equation}
  \label{eq:phifourier}
  \phi(\bm r)=\sum_{\bm G}\phi_{\bm G}(z)e^{i\bm G\cdot\bm r_\|}
\end{equation}
where $\phi_{\bm G}(z)$ is the 2D Fourier coefficient of the potential at
the height $z$. Substitution in Eq. \eqref{eq:poissonRed} yields the
ordinary differential equation
 \begin{eqnarray}
 \frac{d^2}{dz^2}\phi_{\bm G}(z)-G^2\phi_G(z)=
   -\frac{4 \pi}{A} \delta^{(1)}(z),
 \end{eqnarray}
for each coefficient, an homogeneous equation for $z\ne0$ which may be trivially
solved. After applying boundary conditions at $z=0$ and regularity
conditions at infinity (for $G\ne 0$), we obtain
\begin{equation}
  \label{eq:phiG}
  \phi_{\bm G}(z)=
  \begin{cases}
    -\dfrac{2\pi}{A}\abs{z}&\text{if }G=0,\\
    \dfrac{2 \pi}{AG}  e^{- G \abs{z}}&\text{if }G\neq 0.
  \end{cases}
\end{equation}
The case $G=0$ corresponds to the potential produced by a uniformly charged plane,
while the case $G\ne 0$ is the potential produced by a sinusoidal
charge density on a plane, given by the
symmetric solution that decays exponentially as we get away, upwards
or downwards, from
the source plane. Substituting into Eq. \eqref{eq:phifourier} yields
\begin{eqnarray}
  \label{eq:phiG1}
 \phi(\bm r) = -\frac{2 \pi}{A}  \abs{z} +{\sum_{\bm G}}' \frac{2 \pi
   }{AG} e^{- G \abs{z}} e^{i \bm G\cdot\bm r_\|},
 \end{eqnarray}
where the prime indicates that the term $G=0$ should be ommited from
the sum.

We consider now the basis of our 2D lattice, with equal charges at the
four positions $(\pm x_0, \pm y_0,0)$. Each of the corresponding four
sublattices produces a potential as that in Eq. \eqref{eq:phiG1} but
shifted by the corresponding basis vector, yielding
\begin{equation}
  \label{eq:phiG4}
  \phi(\bm r) = -\frac{8\pi}{A} \abs{z}
    +{\sum_{\bm G}}' \frac{8\pi}{AG} e^{- G \abs{z}}\cos(G_x x_0)
    \cos(G_y y_0) e^{i \bm G\cdot\bm r_\|}.
\end{equation}
Finally, we shift the origin of Eq. \eqref{eq:phiG4} vertically to each of the planes shown in
Fig. \ref{fig:sideview} and multiply by the corresponding charge $q_0$
or $-q_0$ to obtain the contribution to the total potential. The charge $q_0$ corresponds to
the planes with heights of the form $z_0+4nc$ and $-z_0+(4n+2)c$, with
$n=\ldots-2,-1,0,1,2\ldots$ an arbitrary integer, while the charge
$-q_0$ corresponds to the planes with heights $z_0+(4n+2)c$ and
$-z_0+4(n+1)c$.

@@

  \caption{(a): Schematic experimental set-up for
    the fabrication of GRIN PS structures. We show a side view of the
    electrochemical cell with the shape of a
    rectangular prism. We indicate the Si wafer (dark
    gray), the teflon cell (light gray), the sealing o-ring (black
    circles), the insulated feeding cable (white lines within black
    lines), the uncovered platinum tip (white) within the
    electrolyte (blue), the electrical contact below the sample
    (yellow) and the controlled current source. We indicate some geometrical
    parameters of the arrangement $x_0$, $z_0$, $a$ and $c$
    and a few schematic current flux
    lines. (b) and (c): Top and side schematic views of expected resulting sample
    illustrating eight regions $1\ldots8$ of decreasing porosity and thickness,
    illustrating the changes in pore radii,
    length, and density and the porosity gradient.}
  \label{fig:DE1}
\end{figure}

@@

\subsection{Fabrication of PS  GRIN Single Layers}
\label{sec:fabrication-ps-grin}
Single layer PS structures were fabricated through electrochemical
anodization of (100) oriented $p$-type crystalline Si wafer
(resistivity 0.002 - 0.005 $ \Omega\,\text{cm}$), under galvanostatic
conditions. The process was performed at room temperatured with an
electrolytic mixture of aqueous hydrofluoric
acid (HF) (concentration 48\% wt), glycerol (purity 99.8 \% wt) and
ethanol (purity 99.9\% wt) in a 3:7:1 volume ratio.
After the anodizing process, the samples were rinsed with
ethanol (purity: 99.9 $ \% $  wt). The electrolytic cell had the shape
of a rectangular prism as shown in Fig. \ref{fig:DE1}, with a base of
sides $a=2.01\,\text{cm}$ and $b=1.48\,\text{cm}$, with an uncovered sample area
of about $3\,\text{cm}^2$. The height of the liquid was set at
$c=1\,\text{cm}$. The cathode was a
platinum wire with a diameter of $0.41\,\text{mm}$ insulated with
teflon tape, with an uncovered
tip located at $\bm r_0=(x_0,
y_0,z_0)=(0.2\,\text{cm},0.74\,\text{cm},0.9\,\text{cm})$.
We manufactured three samples $S_1$, $S_2$ and $S_3$ by applying a
current $I_1=5\,\text{mA}$, $I_2=10\,\text{mA}$ and
$I_3=20\,\text{mA}$, respectively, during a time of $t=250s$.

\subsection{Reflectance Measurement}
\label{sec:refl-meas}
We measured the absolute reflectance at different positions over our
samples using Universal reflectance accessory  equipped {\em Perkin Elmer Lambda 950}
ultraviolet-visible-near infrared (UV-Vis-NIR) spectrophotometer, with a rectangular spot size of
$0.7\text{mm}\times2.0\text{mm}$, a slit width of $0.6\text{mm}$ in
a wavelength range from $300\text{nm}$ to
$1400\text{nm}$.

For interpretation of the data we used
the standard formula \cite{stenzel2015physics} for the reflectance of
a three media system:
an air environment  (0), the PS film (1) and the crystalline Si (c-Si) substrate (2),
\begin{equation}\label{Eq:ECMR}
  R=\left |\frac{r _ {_ {01}}r _ {_ {12}} e^{2i\psi}}{1+r _ {_
        {01}}r _ {_ {12}} e^{2i\psi}}\right |^2
\end{equation}
where $r_{01}$ and $r_{12}$ are the Fresnel reflectance
coefficients of the air/PS and PS/c-Si interfaces \cite{sattler2017silicon}
and $\psi=k_1^\perp d$, where $k_1^\perp$ is
the component of the wavevector perpendicular to the surface within
the porous layer of thickness $d$. The reflectance depends on the
thickness of the film and on the index of refraction $n_1=n_{\text{PS}}$ of the porous
layer, which we relate to the porosity through Bruggeman's
effective medium theory \cite{theiss1997optical},
\begin{equation}\label{Eq:Brugg}
 p\frac{1-\epsilon_1} {1+\epsilon_1} + (1-p)
  \frac{\epsilon_2 - \epsilon_1}
  {\epsilon_2+\epsilon_1}=0,
\end{equation}
where $\epsilon_2=\epsilon_{\text{Si}}$ is the dielectric function of
the c-Si substrate, and
$\epsilon_1=n_{\text{PS}}^2$ is the dielectric function of the PS layer.
We fitted the reflectance spectra using the porosity $p$ and the
thickness $d$ and a scale calibration factor $s$ as fitting
parameters, and thus, we obtained the porosity
$p$ and the etching rate $v=d/t$ for different values of the local
current density $j_\perp$ expected at several positions on different
samples.

@@

\section{Discussion and Results}
\label{sec:discussion-results}
\subsection{Current Density}
\label{sec:current-density}
\begin{figure}
  \centering
  \begin{picture}(400,300)
    \put(0,0){\includegraphics[width=.8\textwidth]{j_Ivsx.png}}
    \put(160,140){\includegraphics[trim=50 50 70 80,clip,width=.4\textwidth]{j_Ivsxy.png}}
  \end{picture}
  \caption{Normalized current density $j_\perp(x,y,0)/I$ for a system
    as in Figs. \ref{f:celda} and \ref{fig:DE1} for a cell of
    length $a = 2.01\,\text{cm}$ and width $b=1.48\,\text{cm}$, with
    the surface of the electrolyte at a heigth $c=1.0\,\text{cm}$, with a
    point-like electrode at $\bm r_0=(0.2\,\text{cm},
    0.74\,\text{cm},0.9\,\text{cm})$,  calculated through
    Eq. (\ref{Eq:J}) along the center line
    $y=0.74\,\text{cm}$. Inset: current over the whole surface.}
  \label{fig:DR1}
\end{figure}
In Fig. \ref{fig:DR1} we show the normalized current density $j_\perp(x,y,0)/I$ calculated through
Eq. \eqref{Eq:J} at the surface of the sample for our cell, as shown
in Figs. \ref{f:celda} and \ref{fig:DE1} and described in
Sect. \ref{sec:fabrication-ps-grin}). For these parameters we obtained
a large range of values for $ j_\perp/I\sim 0.11-.72\,\text{cm}^{-2}$
that vary mostly along the $x$ direction. Of course, this would differ
for cells with different aspect ratios and for different positions of
the electrode.

\subsection{Samples }
\label{sec:ps-grin-single}
\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{G3I20proc}
  \caption{ Photograph of sample $S_3$ prepared as described in
    Sec. \ref{sec:fabrication-ps-grin} using the same parameters as
    in Fig. \ref{fig:DR1} with a current $ I_3=20 \text{mA}$,
    during a time $t=250\text{s}$. We indicate the centerline with a
    dashed line and eight regions $x_i$
    $(i=1, 2,...8)$ centered at positions,
    $x'_1=0.2\text{cm}$,  $x'_2=0.4\text{cm}$, $x'_3=0.6\text{cm}$,
    $x'_4=0.8\text{cm}$, $x'_5=1\text{cm}$, $x'_6=1.2\text{cm}$,
    $x'_7=1.4\text{cm}$, $x'_8=1.6\text{cm}$, with $y_n=0.74\,\text{cm}$
    on which we measured reflectance spectra. Here,
    $x'_n=x_n+0.14\,\text{cm}$ is the position from the left side sealing o-ring.}
  \label{fig:SI3}
\end{figure}

>>>>

  non-homogeneous multilayered devices with lateral gradients for

@@

\section{Introduction}
\label{sec:introduction}
Surfaces with a gradient in the refractive index (GRIN) are of
importance as they allow the engineering and tuning of the optical
phase along their surface, yielding applications such as flat
lenses. Furthermore, there are important biological applications
which combine topographical features with index of refraction
gradients \cite{pierscionek2012gradient}. Preparation of structures
with gradients in other properties have also shown their
usefulness. As an example, the
roughness dependent  response of cells has been tested on
materials used for applications in medical implants
\cite{bai2020bioinspired}.
Gradients have also been employed for several
technological applications such as high-yield screening of catalysts,
sensing materials and bio-molecules \cite{maier2007combinatorial, jayaraman2004construction,
potyrailo2008combinatorial}
Apart from some methods developed by S.E. Fosdick et al. \cite{fosdick2010two},
%[JACS  2010, 132, 9226-9227],
synthesis of  Ag-Au alloy gradients on steel and chemical composition
gradients of CdS layers
on gold electrodes have also been obtained \cite{wang2019fabrication}.
Among electrochemical methods, changes in the composition and doping density
of  conducting polymers have been used to produce gradients using Indium Tin
Oxide electrodes \cite{inagi2010bipolar}.
Electrochemically induced potential gradients have been shown to catalyze the
reactions and gradient doping of polymers \cite{qin2020bipolar, ishiguro2011gradient}.
As compared to the above mentioned techniques, a relatively economical,
fast and easy approach for the fabrication
of gradient materials has been the use  of asymmetrical
electrode configurations in the electrochemical synthesis of porous silicon \cite{sailor2012porous}.
This technique has the additional advantage of being compatible
with very-large-scale-integration (VLSI) devices that may be
integrated in  microelectronic circuits.

Although porous silicon (PS) initially stimulated the interest of the scientific
community due to its photo-luminescence in the visible range at room
temperature, it has presently
been recognized as a multifaceted optical material due to its large surface
area, bio-compatibility, ease of fabrication and tunable refractive index.
Applications based on porous silicon now cover various fields such as
chemical sensors and biosensors \cite{kumar2020porous},
microelectronics and micro-mechanical systems (MEMS)
\cite{martinez2016dual}, as well as a
range of optical \cite{estevez2014porous} and opto-electronic applications
\cite{ramadan2020fabrication}. Specifically,
the temporal variation of the current density results in a variation of
porosity along the depth, allowing the easy fabrication of different kinds
\cite{perez2018reflectivity} of 1D dielectric multilayered
structures.
Although the conventional fabrication of PS yields laterally
homogeneous samples, with invariant structural and optical properties
along the surface, another configuration where a Pt pin electrode is placed
perpendicular to the silicon wafer, produces porous
silicon structures with a lateral gradient in properties such as
thickness, pore size and density, and thus, refractive index and
optical thickness \cite{khung2008using, clements2011mesenchymal}.
The resulting porous surface can have pore sizes ranging from a few nanometers
to  few micrometers  \cite{wang2015screening}. The range of the dimensions of the pores and
the corresponding  thickness range on the same chip can be controlled by
adjusting the location of the electrode, the  anodizing current
and the etching time. Different
applications have been found for the resulting structures
\cite{collins2002determining}. For example, Sailor's group
\cite{sailor2012porous, collins2002determining} applied it
for the determination of protein size.
Additionally, a variation in pH of the solution could gate the
trapping/release of bio-molecules at regions with different porosity, a
result potentially useful for drug delivery
applications given the bio-compatible nature of PS.
A similar asymmetric electrode configuration was used to fabricate
multilayered optical filters with lateral a gradient
\cite{li2005painting} and
for developing ethanol sensors \cite{sailor2012porous}.
Additionally, it has been shown that the combination of thermal  oxidation  and
infiltration of TiO$_2$  by atomic layer deposition allows the manufacture
of transparent GRIN
optical elements with a high refractive index contrast \cite{ocier2017tunable}.
Kruger et al. \cite{krueger2018electrochemical} used
a porous silicon 1D GRIN structure infiltrated with polymer to
synthesize flat micro-lens arrays.
Recently, J Wang et al. has shown the fabrication of
a miniature spectrometer with a PS based rugate filter using a
radial interfacial potential distribution \cite{wang2019fabrication}
On the other hand, the effect of different topographical
features present on the same chip, has also been
used to study the dependence of the adhesion of certain cells on the
surface topography \cite{khung2008using}.

@@

with that within an infinite fictitious space  empty but for the point charge

@@

The electrostatic potential produced by an infinite array of
charges expressed as a sum in real space of Coulomb terms has ill
convergence properties. The potential can be obtained
using the Ewald summation technique
\cite{ewald1921calculation, belhadj1991molecular-dynamics,
  schmidt1997multipole, grzybowski2000ewald, toukmaji2000efficient,
  brodka2004ewald, campione2012ewald}, which yields
two rapidly converging series, one in real space and another one in reciprocal space.
Nevertheless, as we need the potential and the field to obtain the
current density on the surface of the sample on which  no
real nor image charges lie, we may perform a somewhat simpler plane-wise
summation, where we first obtain the potential and field produced by a
simple periodic lattice of charges using a fast convergent sum over
the two dimensional (2D)
reciprocal vectors \cite{nijboer1957calculation, nijboer1958internal,
  de_wette1965internal, watson1981madelung, grindlay1981k},
then sum over the charges that form the basis
of the 2D {\em crystal} plane, and finally sum over all the
planes.

We consider first a system made of identical {\em unit} charges
occupying the positions $\bm R$ of a 2D Bravais lattice $\{\bm R\}$,
which we will take as a rectangular lattice with lattice parameters
$2a$ and $2b$, and lying, for the time being, at the $z=0$ plane.
The potential $\phi$ it produces obeys
\begin{equation}
  \label{eq:poissonRed}
  \nabla^2\phi(\bm r)=-4\pi\sum_{\bm R}\delta^{(2)}(\bm r_\|-\bm
  R)\delta^{(1)}(z)=-\frac{4\pi}{A}\sum_{\bm G}e^{i\bm G\cdot\bm r_\|}\delta^{(1)}(z),
\end{equation}
where $\bm r_\|$ is the projection of the observation position $\bm
r=(x,y,z)$ onto the $xy$ plane, $\{\bm G\}$ is the 2D
reciprocal lattice defined through $e^{i\bm G\cdot\bm R}=1$ for all
$\bm R$ and $\bm G$, $A=4ab$ is the area of the unit cell
and we used the Fourier representation of the 2D
periodically repeated delta function. We introduce a 2D Fourier representation for
the potential
\begin{equation}
  \label{eq:phifourier}
  \phi(\bm r)=\sum_{\bm G}\phi_{\bm G}(z)e^{i\bm G\cdot\bm r_\|}
\end{equation}
where $\phi_{\bm G}(z)$ is the 2D Fourier coefficient of the potential at
the height $z$. Substitution in Eq. \eqref{eq:poissonRed} yields the
ordinary differential equation
 \begin{eqnarray}
 \frac{d^2}{dz^2}\phi_{\bm G}(z)-G^2\phi_G(z)=
   -\frac{4 \pi}{A} \delta^{(1)}(z),
 \end{eqnarray}
for each coefficient, an homogeneous equation for $z\ne0$ which may be trivially
solved. After applying boundary conditions at $z=0$ and regularity
conditions at infinity (for $G\ne 0$), we obtain
\begin{equation}
  \label{eq:phiG}
  \phi_{\bm G}(z)=
  \begin{cases}
    -\dfrac{2\pi}{A}\abs{z}&\text{if }G=0,\\
    \dfrac{2 \pi}{AG}  e^{- G \abs{z}}&\text{if }G\neq 0.
  \end{cases}
\end{equation}
The case $G=0$ corresponds to the potential produced by a uniformly charged plane,
while the case $G\ne 0$ is the potential produced by a sinusoidal
charge density on a plane, given by the
symmetric solution that decays exponentially as we get away, upwards
or downwards, from
the source plane. Substituting into Eq. \eqref{eq:phifourier} yields
\begin{eqnarray}
  \label{eq:phiG1}
 \phi(\bm r) = -\frac{2 \pi}{A}  \abs{z} +{\sum_{\bm G}}' \frac{2 \pi
   }{AG} e^{- G \abs{z}} e^{i \bm G\cdot\bm r_\|},
 \end{eqnarray}
where the prime indicates that the term $G=0$ should be omitted from
the sum.

We consider now the basis of our 2D lattice, with equal charges at the
four positions $(\pm x_0, \pm y_0,0)$. Each of the corresponding four
sub-lattices produces a potential as that in Eq. \eqref{eq:phiG1} but
shifted by the corresponding basis vector, yielding
\begin{equation}
  \label{eq:phiG4}
  \phi(\bm r) = -\frac{8\pi}{A} \abs{z}
    +{\sum_{\bm G}}' \frac{8\pi}{AG} e^{- G \abs{z}}\cos(G_x x_0)
    \cos(G_y y_0) e^{i \bm G\cdot\bm r_\|}.
\end{equation}
Finally, we shift the origin of Eq. \eqref{eq:phiG4} vertically to each of the planes shown in
Fig. \ref{fig:sideview} and multiply by the corresponding charge $q_0$
or $-q_0$ to obtain the contribution to the total potential. The charge $q_0$ corresponds to
the planes with heights of the form $z_0+4nc$ and $-z_0+(4n+2)c$, with
$n=\ldots-2,-1,0,1,2\ldots$ an arbitrary integer, while the charge
$-q_0$ corresponds to the planes with heights $z_0+(4n+2)c$ and
$-z_0+4(n+1)c$.

@@

  \caption{(a): Schematic experimental set-up for
    the fabrication of GRIN PS structures. We show a side view of the
    electrochemical cell with the shape of a
    rectangular prism. We indicate the Si wafer (dark
    gray), the Teflon cell (light gray), the sealing o-ring (black
    circles), the insulated feeding cable (white lines within black
    lines), the uncovered platinum tip (white) within the
    electrolyte (blue), the electrical contact below the sample
    (yellow) and the controlled current source. We indicate some geometrical
    parameters of the arrangement $x_0$, $z_0$, $a$ and $c$
    and a few schematic current flux
    lines. (b) and (c): Top and side schematic views of expected resulting sample
    illustrating eight regions $1\ldots8$ of decreasing porosity and thickness,
    illustrating the changes in pore radii,
    length, and density and the porosity gradient.}
  \label{fig:DE1}
\end{figure}

@@

\subsection{Fabrication of PS  GRIN Single Layers}
\label{sec:fabrication-ps-grin}
Single layer PS structures were fabricated through electrochemical
anodization of (100) oriented $p$-type crystalline Si wafer
(resistivity 0.002 - 0.005 $ \Omega\,\text{cm}$), under galvanostatic
conditions. The process was performed at room temperature with an
electrolytic mixture of aqueous hydrofluoric
acid (HF) (concentration 48\% wt), glycerol (purity 99.8 \% wt) and
ethanol (purity 99.9\% wt) in a 3:7:1 volume ratio.
After the anodizing process, the samples were rinsed with
ethanol (purity: 99.9 $ \% $  wt). The electrolytic cell had the shape
of a rectangular prism as shown in Fig. \ref{fig:DE1}, with a base of
sides $a=2.01\,\text{cm}$ and $b=1.48\,\text{cm}$, with an uncovered sample area
of about $3\,\text{cm}^2$. The height of the liquid was set at
$c=1\,\text{cm}$. The cathode was a
platinum wire with a diameter of $0.41\,\text{mm}$ insulated with
Teflon tape, with an uncovered
tip located at $\bm r_0=(x_0,
y_0,z_0)=(0.2\,\text{cm},0.74\,\text{cm},0.9\,\text{cm})$.
We manufactured three samples $S_1$, $S_2$ and $S_3$ by applying a
current $I_1=5\,\text{mA}$, $I_2=10\,\text{mA}$ and
$I_3=20\,\text{mA}$, respectively, during a time of $t=250s$.

\subsection{Reflectance Measurement}
\label{sec:refl-meas}
We measured the absolute reflectance at different positions over our
samples using Universal reflectance accessory  equipped {\em Perkin Elmer Lambda 950}
ultraviolet-visible-near infrared (UV-Vis-NIR) spectro-photometer, with a rectangular spot size of
$0.7\text{mm}\times2.0\text{mm}$, a slit width of $0.6\text{mm}$ in
a wavelength range from $300\text{nm}$ to
$1400\text{nm}$.

For interpretation of the data we used
the standard formula \cite{stenzel2015physics} for the reflectance of
a three media system:
an air environment  (0), the PS film (1) and the crystalline Si (c-Si) substrate (2),
\begin{equation}\label{Eq:ECMR}
  R=\left |\frac{r _ {_ {01}}r _ {_ {12}} e^{2i\psi}}{1+r _ {_
        {01}}r _ {_ {12}} e^{2i\psi}}\right |^2
\end{equation}
where $r_{01}$ and $r_{12}$ are the Fresnel reflectance
coefficients of the air/PS and PS/c-Si interfaces \cite{sattler2017silicon}
and $\psi=k_1^\perp d$, where $k_1^\perp$ is
the component of the wave-vector perpendicular to the surface within
the porous layer of thickness $d$. The reflectance depends on the
thickness of the film and on the index of refraction $n_1=n_{\text{PS}}$ of the porous
layer, which we relate to the porosity through Bruggeman's
effective medium theory \cite{theiss1997optical},
\begin{equation}\label{Eq:Brugg}
 p\frac{1-\epsilon_1} {1+\epsilon_1} + (1-p)
  \frac{\epsilon_2 - \epsilon_1}
  {\epsilon_2+\epsilon_1}=0,
\end{equation}
where $\epsilon_2=\epsilon_{\text{Si}}$ is the dielectric function of
the c-Si substrate and
$\epsilon_1=n_{\text{PS}}^2$ is the dielectric function of the PS layer.
We fitted the reflectance spectra using the porosity $p$ and the
thickness $d$ and a scale calibration factor $s$ as fitting
parameters, and thus, we obtained the porosity
$p$ and the etching rate $v=d/t$ for different values of the local
current density $j_\perp$ expected at several positions on different
samples.

@@

\section{Results and Discussion}
\label{sec:discussion-results}
\subsection{Current Density}
\label{sec:current-density}
\begin{figure}
  \centering
  \begin{picture}(400,300)
    \put(0,0){\includegraphics[width=.8\textwidth]{j_Ivsx.png}}
    \put(160,140){\includegraphics[trim=50 50 70 80,clip,width=.4\textwidth]{j_Ivsxy.png}}
  \end{picture}
  \caption{Normalized current density $j_\perp(x,y,0)/I$ for a system
    as in Figs. \ref{f:celda} and \ref{fig:DE1} for a cell of
    length $a = 2.01\,\text{cm}$ and width $b=1.48\,\text{cm}$, with
    the surface of the electrolyte at a height $c=1.0\,\text{cm}$, with a
    point-like electrode at $\bm r_0=(0.2\,\text{cm},
    0.74\,\text{cm},0.9\,\text{cm})$,  calculated through
    Eq. (\ref{Eq:J}) along the center line
    $y=0.74\,\text{cm}$. Inset: current over the whole surface.}
  \label{fig:DR1}
\end{figure}
In Fig. \ref{fig:DR1} we show the normalized current density $j_\perp(x,y,0)/I$ calculated through
Eq. \eqref{Eq:J} at the surface of the sample for our cell, as shown
in Figs. \ref{f:celda} and \ref{fig:DE1} and described in
Sect. \ref{sec:fabrication-ps-grin}). For these parameters we obtained
a large range of values for $ j_\perp/I\sim 0.11-.72\,\text{cm}^{-2}$
that vary mostly along the $x$ direction. Of course, this would differ
for cells with different aspect ratios and for different positions of
the electrode.

\subsection{Samples }
\label{sec:ps-grin-single}
\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{G3I20proc}
  \caption{ Photograph of sample $S_3$ prepared as described in
    Sec. \ref{sec:fabrication-ps-grin} using the same parameters as
    in Fig. \ref{fig:DR1} with a current $ I_3=20 \text{mA}$,
    during a time $t=250\text{s}$. We indicate the center-line with a
    dashed line and eight regions $x_i$
    $(i=1, 2,...8)$ centered at positions,
    $x'_1=0.2\text{cm}$,  $x'_2=0.4\text{cm}$, $x'_3=0.6\text{cm}$,
    $x'_4=0.8\text{cm}$, $x'_5=1\text{cm}$, $x'_6=1.2\text{cm}$,
    $x'_7=1.4\text{cm}$, $x'_8=1.6\text{cm}$, with $y_n=0.74\,\text{cm}$
    on which we measured reflectance spectra. Here,
    $x'_n=x_n+0.14\,\text{cm}$ is the position from the left side sealing o-ring.}
  \label{fig:SI3}
\end{figure}

In Fig. \ref{fig:SI3} we show a photograph of one of our samples,
$S_3$
prepared as described in Sec. \ref{sec:fabrication-ps-grin}, with the
same conditions as those corresponding to Fig.~\ref{fig:DR1} applying
a current
$I_3=20\,\text{mA}$ during a time $t=250\,s$. We notice a series of
visible interference fringes that get wider as we move towards the
right end of the sample, consistent with a inhomogeneous film that
gets thinner towards the right, and qualitatively consistent with
Fig. \ref{fig:DR1} that shows a current density that decays towards
the right. Equal color lines seem to correspond to iso-current lines of
Fig. \ref{fig:SI3}, except very close to the the walls of the cell,
where the film rapidly becomes very narrow. The reason is that there
is a small region below the cell walls where the electrolyte penetrates
up to the sealing o-ring (see upper panel of
Fig. \ref{fig:DE1}) and towards which the current leaks from the
neighborhood of the wall. The width of this region is
$w=0.14\,\text{cm}$. Thus, we may expect our formula \eqref{Eq:J}
to {\em fail} within that region and close to the border, at distances of the order of the radius
of the cross-section of the o-ring, about a millimeter. The
figure shows eight positions
$x'_n=0.2\,\text{cm},0.4\,\text{cm},\ldots,1.6\,\text{cm}$ on which we
measured the near normal incidence reflection spectra of the
sample. Here, we defined $x'_n=x_n+w$ as the distance to the edge of
the etched sample, which corresponds to the position of the sealing o-ring,
whereas $x_n$ is the corresponding distance to the wall of the
cell.
